Learning Network Size while Training with ShrinkNets

### 2018-02-09 paper/ICML/impl.tex
edits

In this section we discuss practical aspects of ShrinkNets.

@@

Switch layers are initialized with weights sampled from $\mathcal{N}(0,1)$, and
their values change as part of the training process so as to swith \emph{on} or
{off} neurons. When neurons are deactivated by switch layers, we want to remove
them to shrink the network size. Because after removing a neuron it will no
longer contribute to improving prediction accuracy, we must devise strategies
that remove neurons while being robust to changing values. We observe three
strategies:

@@

\textbf{Sign change strategy}: Under this strategy neurons are removed when the
weight value changes its sign. This strategy works well in practice but it is
also sensible to gradient noise. \ra{explain what's the consequence of that} If
we sample a gradient that is not representative of the dataset when performing
gradient descent then we might wrongly kill a neuron.\ra{last sentence is
unclear}

\textbf{Sign variance strategy}: We measure the exponential moving variance of
the sign ($-1, 1$) of each parameter in the \textit{switch layer}. When the
value of the exponential moving variances goes over a predefined threshold, we
consider the parameter is deactivating the neuron, and therefore we remove it.
This strategy introduces two extra parameters, one to control the behavior of
the inertia of the variance, and the threshold, but we have found it is the
best performing in practice.

The last two strategies perform the best in practice; they reduce the network
parameters by an order of magnitude with only a minimal impact on the final
error obtained by the model.

@@

\noindent\textbf{Preparing for Inference: } With ShrinkNets we obtain
reduced-sized networks while training, which is the first benefit towards faster
inference. This networks are readily available for inference. However, because
they include switch layers---and therefore more parameters---they introduce
unnecessary overhead at inference time. To avoid this overhead, we reduce the
network parameters by combining each switch layer with its respective network
layer by multiplying the respective parameters.  \ra{can't we just simply
prevent the following problem at network design time?}In the unlikely case that a switch
layer is sandwiched between two non-linearly scalable layers, we leave the
switch layer as is.

\noindent\textbf{Neural Garbage Collection: }ShrinkNets decide on-the-fly which
neurons to remove. Since ShrinkNets remove a large fraction of neurons, we must
dynamically resize the network at runtime to not carry on the 
unnecessary overhead. We implement a neural garbage collection method as part of
our library which takes care of updating the necessary network layers as well as
updating optimizer state to reflect the neuron removal.

>>>>

In this section we discuss practical aspects of ShrinkNets, including the
neuron removal  process and several additional optimizations.

@@

Switch layers are initialized with weights sampled from $\mathcal{N}(0,1)$;
their values change as part of the training process so as to switch \emph{on}
or \emph{off} neurons. When neurons are deactivated by switch layers, we need
to remove them to actually shrink the network size. Since after removing a
neuron it will no longer be able to contribute to  prediction accuracy, we
must devise strategies that remove neurons while remaining robust to changing
values. We observe three strategies:

@@

\textbf{Sign change strategy}: Under this strategy neurons are removed when
the weight value changes its sign. This strategy works well in practice but it
is also sensitive to gradient noise. \ra{explain what's the consequence of
that} If we sample the gradient in a way that is not fully representative of
the dataset we might  experience one-time zero crossing which could wrongly
kill a neuron.

\textbf{Sign variance strategy}: Instead of killing  on the first zero
crossing, we can instead measure the exponential moving variance of the sign
($-1, 1$) of each parameter in the \textit{switch layer}. When the value of
the exponential moving variances goes over a predefined threshold, we consider
the neuron deactivated, and therefore we remove it. This strategy introduces
two extra parameters, one to control the behavior of the inertia of the
variance, and the threshold, but we have found it is the best performing in
practice.

The last two strategies perform better than the threshold strategy.  Both
effectively allow us to shrink the network by varying $\lambda$, and result in
networks that are both small and perform well.

@@

\noindent\textbf{Preparing for Inference: } With ShrinkNets we obtain reduced-
sized networks during training, which is the first steps towards faster
inference. This networks are readily available for inference. However, because
they include switch layers---and therefore more parameters---they introduce
unnecessary overhead at inference time. To avoid this overhead, we reduce the
network parameters by combining each switch layer with its respective network
layer by multiplying the respective parameters before emitting the final
trained network.   \ra{can't we just simply prevent the following problem at
network design time?}In the unlikely case that a switch layer is sandwiched
between two non-linearly scalable layers, we leave the switch layer as is.

\noindent\textbf{Neural Garbage Collection: }ShrinkNets decides on-the-fly
which neurons to remove. Since ShrinkNets removes a large fraction of neurons,
we must dynamically resize the network at runtime to not unnecessarily impact
network training time. We implemented a neural garbage collection method as
part of our library which takes care of updating the necessary network layers
as well as updating optimizer state to reflect the neuron removal.

### 2018-02-09 paper/ICML/theory.tex

%\subsection{Relation to Group Sparsity}
\noindent\textbf{Relation to Group Sparsity (LASSO): } ShrinkNets removes neurons,
i.e., inputs and outputs of layers. For a fully connected layer defined as:

\begin{equation} \label{fully_connected}
  f_{\bm{A}, \bm{b}}(\bm{x})=a(\bm{Ax + b})
\end{equation}

Removing an input neuron $j$ is equivalent to have $\left(\bm{A}^T\right)_j =
\bm{0}$. Removing an output neuron $i$ is the same as having $\bm{A}_i = \bm{0}$
and $\bm{b}_i = 0$. Solving optimization problems while trying to set entire
group of parameters to zero is the goal of group sparsity regularization
\cite{}. For any partitioning of the set of parameter defining a model in $p$
groups: $\bm{\theta} = \bigcup_{i=1}^P \bm{\theta}_i$ we define it the following
way:

\begin{equation}
  \Omega_\lambda^{gp} = \lambda \sum_{i=1}^p \sqrt{\#\bm{\theta_i}} \norm{\bm{\theta_i}}_2 \\
\end{equation}

In fully-connected layers, the groups are either: columns of
$\bm{A}$ if we want to remove inputs, or rows of $\bm{A}$ and the corresponding
entry in $\bm{b}$ if we want to remove outputs. For simplicity, we focus
our analysis in the simple one-layer case. In this case, filtering outputs does
not make sense, this is why we will only consider the former case. The
group sparsity regularization then becomes:

\begin{equation} \label{group_sparsity_regularization}
  \Omega_\lambda^{gp} = \lambda \sum_{j=1}^p \norm{\bm{\left(A^T\right)_j}}_2 \\
\end{equation}

Because $\forall i, \#\bm{\theta}_i = n$, \ra{is \# the general way of expressing
cardinality? why not |x|?}to make the notation simpler, we
embedded $\sqrt{n}$ inside $\lambda$.

Group sparsity and ShrinkNets try to achieve the same goal. We discuss next how
they are related to each other. First let's recall the two problems. The
original ShrinkNet problem is:

\begin{equation}
  \min_{\bm{A}, \bm{\beta}} \norm{\bm{y} - \bm{A}\diag{\bm{\beta}}\bm{x}}_2^2 + \lambda \norm{\bm{\beta}}_1
\end{equation}

And the Group Sparsity problem is:

\begin{equation}
  \min_{\bm{A}} \norm{\bm{y} - \bm{A}\bm{x}}_2^2 + \Omega_\lambda^{gp}
\end{equation}

We can prove that under the condition: $\forall j\in \intint{1, p},
\norm{\left(\bm{A}^T\right)_j}_2 = 1$ the two problems are equivalent
(\cref{gps_equivalence}). However if we relax this constraint then shrinknet
becomes non-convex and has no global minimum
(\cref{unconstrained_non_convex,unconstrained_shrinknet_no_min}). Fortunately,
by adding an extra term to the ShrinkNet regularization term we can proove that:

\begin{equation}
  \min_{\bm{A}, \bm{\beta}} \norm{\bm{y} - \bm{A}\diag{\bm{\beta}}\bm{x}}_2^2 + \Omega_\lambda^s + \lambda_2\norm{A}_p^p
\end{equation}

has global minimums for all $p>0$ (\cref{shrinknet_regularized_minimum}).
This is the reason we defined the \textit{regularized ShrinkNet penalty} earlier
this way:\ra{the notation is slightly different, better to make it consistent}

\begin{equation}
  \Omega_{\lambda, \lambda_2, p}^{rs} = \lambda\norm{\bm{\beta}}_1 + \lambda_2\norm{\bm\theta}_p^p
\end{equation}

In practice we observed that $p=2$ or $p=1$ are good choice, while the latter
will also introduce additional sparsity in the parameters.

>>>>

%!TEX root=paper.tex
%\subsection{Relation to Group Sparsity}
\noindent\textbf{Relation to Group Sparsity (LASSO): } ShrinkNets removes neurons,
i.e., inputs and outputs of layers. For a fully connected layer defined as:
%
\begin{equation} \label{fully_connected}
  f_{\bm{A}, \bm{b}}(\bm{x})=a(\bm{Ax + b})
\end{equation}
%
removing an input neuron $j$ is equivalent to having $\left(\bm{A}^T\right)_j =
\bm{0}$. Removing an output neuron $i$ is the same as setting $\bm{A}_i = \bm{0}$
and $\bm{b}_i = 0$. Solving optimization problems while trying to set entire
group of parameters to zero is the goal of group sparsity regularization
\cite{xxx}. In  any partitioning of the set of parameters $\theta$ defining a model in $p$
groups: $\bm{\theta} = \bigcup_{i=1}^P \bm{\theta}_i$, group sparsity is defined as: \srm{you used to say ``we define `it' as:'' -- I am assume `it' is group sparsity}
%
\begin{equation}
  \Omega_\lambda^{gp} = \lambda \sum_{i=1}^p \sqrt{\#\bm{\theta_i}} \norm{\bm{\theta_i}}_2 \\
\end{equation}
%
\srm{define notation -- what is $\lambda$;  what is $\#\theta$, what is $\Omega$? Where does the square root come from?}
In fully-connected layers, the groups are either: columns of
$\bm{A}$ if we want to remove inputs, or rows of $\bm{A}$ and the corresponding
entry in $\bm{b}$ if we want to remove outputs. For simplicity, we focus
our analysis in the simple one-layer case. In this case, filtering outputs does
not make sense, so we only consider removing inputs. The
group sparsity regularization then becomes:
%
\begin{equation} \label{group_sparsity_regularization}
  \Omega_\lambda^{gp} = \lambda \sum_{j=1}^p \norm{\bm{\left(A^T\right)_j}}_2 \\
\end{equation}
%
Because $\forall i, \#\bm{\theta}_i = n$, \ra{is \# the general way of expressing
cardinality? why not $|x|$?}to make the notation simpler, we
embedded $\sqrt{n}$ inside $\lambda$.

Group sparsity and ShrinkNets try to achieve the same goal. We discuss next how
they are related to each other. First let's recall the two problems. The
original ShrinkNet problem is \srm{do you mean ``can be defined formally as'' -- we never gave this definition of ShrinkNets.  Also, you need to explain the formula / notation;  what is $y$, what is $\beta$, what is this equation representing (the optimization function?)}:
%
\begin{equation}
  \min_{\bm{A}, \bm{\beta}} \norm{\bm{y} - \bm{A}\diag{\bm{\beta}}\bm{x}}_2^2 + \lambda \norm{\bm{\beta}}_1
\end{equation}
%
And the Group Sparsity problem is:
%
\begin{equation}
  \min_{\bm{A}} \norm{\bm{y} - \bm{A}\bm{x}}_2^2 + \Omega_\lambda^{gp}
\end{equation}
%

We can prove that under the condition: $\forall j\in \intint{1, p},
\norm{\left(\bm{A}^T\right)_j}_2 = 1$ the two problems are equivalent
(\cref{gps_equivalence}, see Apendix). However if we relax this constraint then ShrinkNets
becomes non-convex and has no global minimum
(\cref{unconstrained_non_convex,unconstrained_shrinknet_no_min}, also in Appendix). Fortunately,
by adding an extra term to the ShrinkNet regularization term we can prove that:
%
\begin{equation}
  \min_{\bm{A}, \bm{\beta}} \norm{\bm{y} - \bm{A}\diag{\bm{\beta}}\bm{x}}_2^2 + \Omega_\lambda^s + \lambda_2\norm{A}_p^p
\end{equation}
%
has global minimums for all $p>0$ (\cref{shrinknet_regularized_minimum}).
This is the reason we defined the \textit{regularized ShrinkNet penalty} above
as \srm{where?}:\ra{the notation is slightly different, better to make it consistent}
%
\begin{equation}
  \Omega_{\lambda, \lambda_2, p}^{rs} = \lambda\norm{\bm{\beta}}_1 + \lambda_2\norm{\bm\theta}_p^p
\end{equation}
%
In practice we observed that $p=2$ or $p=1$ are good a choice; note that the latter
will also introduce additional sparsity into the parameters. \srm{why?}
